Viscous Erosion in Porous Media

### 2018-04-03 docs/methods/erosion_methods.tex
NM

Lastly, we want to briefly examine the number of GMRES iterations required by our simulations since this is typically the most expensive step of the computation. Due to our {\em second-kind} boundary integral formulation, we expect the number of GMRES iterations to be independent of the spatial resolution, $N_{\iin}$. The iteration count, however, may depend on the number of bodies, as well as other factors such as the shapes of the bodies, their configuration, etc. To examine a few of these factors, we show in Table~\ref{itertab} the number of GMRES iterations required to reach a specified tolerance of 1E-8 in several different simulations. The left side of the table shows the results for a 10-body simulation, in which we keep the geometry fixed and vary the spatial resolution $N_{\iin}$. As seen in the table, the iteration count remains essentially fixed as the resolution increases. In the right side of the table, we keep the spatial resolution fixed and vary the number of simulated bodies from 10 to 50. The bodies are all circular with randomly assigned positions and radii (overlapping bodies are avoided). The table shows that the GMRES count, and therefore the computational time, increases roughly linearly with the number of bodies. Thus, with everything else equal, the 50-body simulation from Fig.~\ref{fig:50bodies} is about 50 times more expensive than the single-body simulations.

@@

With the validation complete, we now discuss some of the emergent physics of multi-body erosion revealed by our simulations. We begin with a relatively simple case of 6 eroding bodies, as this will allow us to identify key principles at work in more complex cases (e.g.~the 50-body simulation in Fig.~\ref{fig:50bodies}). Figure~\ref{fig:06bodies} shows the simulated erosion of 6 bodies, each initial circles of various sizes and positions. Over time, erosion not only diminishes the size of the bodies, but also alters their shapes considerable. As in the single-body simulations, corner-like features develop, but not in locations that are easily predictable. Rather, when and where these corners form depends on how neighboring bodies interact as mediated by the Stokes flow.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.9 \textwidth]{./figs/06bod.pdf}
\caption{\label{fig:06bodies} The simulated erosion of 6 bodies with the fixed-pressure drop condition. Erosion causes flat faces to develop wherever bodies were initially very close to one another. This effect leads to the appearance of straight channels between bodies, and causes the bodies themselves to appear distinctly polygonal. The vorticity field (color-bar at right) shows that the flow is concentrated primarily in a single channel that runs horizontally across the domain. }
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1

The initial configuration in  Fig.~\ref{fig:06bodies} was selected to have several points at which the bodies nearly contact one another, for example bodies 1 \& 2, 3 \& 5, and 5 ur \& 6. Intriguingly, erosion appears to remove material preferentially from these near-contact regions, as the narrow spacing between neighboring bodies expands rather rapidly. For example, the spacing between bodies 1 \& 2 is initially very small, but by the second frame, is comparable to the spacing between bodies 3 \& 4. Furthermore, the bodies tend to develop flat faces at the locations of near contact, leading to the appearance of straight channels between bodies. Straight channels can be seen, for example, between bodies 1 \& 2, 3 \& 5, and 5 \& 6. Once formed, these channels persist until the bodies have nearly vanished. Notice that by the final frame, body 4 has vanished completely, and the simulation handles this event without difficulty.

The vorticity field helps not only to visualize regions of high shear, but also to identify the certain preferred channels in which the flow is most intense. As seen in Fig.~\ref{fig:06bodies}, the flow is concentrated primarily in a single channel running horizontally below bodies 1 \& 2 and above bodies 3, 5, \& 6. This main channel widens the fastest, and secondary channels (e.g.~between bodies 1 \& 2 or 3 \& 5 or 5 \& 6) follow. Notice that, in these simulations, the overall vorticity magnitude increases in time as a result of fixing the pressure-drop across the cell.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.95 \textwidth]{./figs/mbodyplots.pdf}
\caption{\label{fig:mbodyplots} plots to go with 6 body simulation.
}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1

>>>>

Lastly, we want to briefly examine the number of GMRES iterations required by our simulations since this is typically the most expensive step. Due to our {\em second-kind} boundary integral formulation, we expect the number of GMRES iterations to be independent of the spatial resolution, $N_{\iin}$. The iteration count, however, may depend on the number of bodies, as well as other factors such as the shapes of the bodies, their configuration, etc. To examine a few of these factors, we show in Table~\ref{itertab} the number of GMRES iterations required to reach a specified tolerance of 1E-8 in several different simulations. The left side of the table shows the results for a 10-body simulation, in which we keep the geometry fixed and vary the spatial resolution $N_{\iin}$. As seen in the table, the iteration count remains essentially fixed as the resolution increases. In the right side of the table, we keep the spatial resolution fixed and vary the number of simulated bodies from 10 to 50. The bodies are all circular with randomly assigned positions and radii (overlapping bodies are avoided). The table shows that the GMRES count, and therefore the computational time, increases roughly linearly with the number of bodies. Thus, with everything else equal, the 50-body simulation from Fig.~\ref{fig:50bodies} is about 50 times more expensive than a single-body simulation.

@@

With the validation complete, we now discuss the emergent physics of multi-body erosion revealed by our simulations. We begin with a relatively simple case of 6 eroding bodies, as this will allow us to identify key principles at work in more complex cases (e.g.~the 50-body simulation in Fig.~\ref{fig:50bodies}). Figure~\ref{fig:06bodies} shows the simulated erosion of 6 bodies, each initial circles of various sizes and positions. Over time, erosion not only diminishes the size of the bodies, but also alters their shapes considerable. As in the single-body simulations, corner-like features develop, but not in locations that are easily predictable. Rather, when and where these corners form depends on how neighboring bodies interact with one another as mediated by the Stokes flow.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.85 \textwidth]{./figs/06bod.pdf}
\caption{\label{fig:06bodies} The simulated erosion of 6 bodies with the fixed-pressure drop condition. Erosion causes flat faces to develop wherever bodies were initially very close to one another. This effect leads to the appearance of straight channels between bodies, and causes the bodies themselves to appear distinctly polygonal. The vorticity field (color-bar at right) shows that the flow is concentrated primarily in a single channel that runs horizontally across the domain. }
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1

The initial configuration in  Fig.~\ref{fig:06bodies} was selected to have several points at which the bodies nearly contact one another, for example between bodies 1 \& 2, 3 \& 5, and 5 \& 6. Intriguingly, erosion appears to remove material preferentially from these near-contact regions, as the narrow spacing between neighboring bodies expands rather quickly. For example, the spacing between bodies 1 \& 2 is initially quite small, but by the second frame, is comparable to the spacing between bodies 3 \& 4. Moreover, the bodies tend to develop flat faces at the locations of near contact, leading to the appearance of straight channels between bodies. Straight channels can be seen, for example, between bodies 1 \& 2, 3 \& 5, and 5 \& 6. Once formed, these channels persist until the bodies have nearly vanished. Notice that by the final frame, body 4 has vanished completely, and the simulation handles this event without difficulty.

The vorticity field helps not only in visualizing regions of high shear, but also in identifying certain preferred channels in which the flow is most intense. Figure~\ref{fig:06bodies} shows the flow to be concentrated primarily within a single channel that runs horizontally below bodies 1 \& 2 and above bodies 3, 5, \& 6. This main channel widens the fastest, and secondary channels (e.g.~between bodies 1 \& 2 or 3 \& 5 or 5 \& 6) follow. Notice that, in these simulations, the overall vorticity magnitude increases in time as a result of fixing the pressure-drop across the cell.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.90 \textwidth]{./figs/mbodyplots.pdf}
\caption{\label{fig:mbodyplots} 
Multi-body simulation results: black curves correspond to the 6-body simulation and gray to the 50-body simulation. (a) The area fraction of the eroding bodies as it changes in time. (b) The incoming flow rate $\umax$ versus time. The fixed-pressure-drop condition causes $\umax$ to increase as the bodies erode. (c) The total drag exerted by all bodies. Due to the fact that $\umax$ increases as the bodies erode, the total drag remains nearly constant and then drops suddenly as the bodies vanish. (d) Normalizing the total drag by the flow rate provides a measurement of the resistance provided by the porous matrix. The resistance decreases rapidly as the bodies erode.
}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1
